\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-28 background.tex
Updated things

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the
\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
the amount of computation required and schedule, $\Delta$ can vary,
hence, in order to satisfy the implicit restriction imposed by the
synchrony hypothesis -- no input event \textit{from the environment} can
be missed -- one needs to calculate the \textit{Worst Case Reaction
  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
between any two consecutive events on any input, else the synchrony
hypothesis is violated. Techniques for calculating the WCRT of
synchronous programs~\cite{boldt07} exist and this analysis is not the
focus of this paper. The opposite of the WCRT, the shortest time for a
reaction, is called the \textit{Best Case Reaction Time} (BCRT), we
denote the WCRT and BCRT in Figure~\ref{fig:2c} using \texttt{W} and
\texttt{B}, respectively. Figure~\ref{fig:2c} shows a continuously
running physical clock (analogous to a clock in digital hardware), the
numerical annotations on the rising edge of the clock mark the logical
ticks from Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e}
for the numbers show the start and end of the logical ticks,
respectively. The end of a logical tick and the start of the next
logical tick happen together. Figure~\ref{fig:2c} shows the mapping of
the logical time to the physical time. For example, the logical tick
\texttt{1} starts at the first rising edge and completes at the second
rising edge of the physical clock, whereas logical tick \texttt{4}
starts at the $4^{th}$ rising edge, but finishes at the $6^{th}$ rising
edge of the physical clock. This difference can be accounted for by the
Java computations required when signal \texttt{A} is input by the
environment (see Figure~\ref{fig:2a}). This elasticity is inherent (and
elegant) to both Esterel and SystemJ.

>>>>

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, execution of every
statement in the logical zero time requires $\delta$ physical time. Time
for a logical tick can thus be summarized as $\Delta = \mathcal{F}
(\delta)$, where $\mathcal{F}$ is the schedule. We call $\Delta$ the
reaction time. Depending upon the amount of computation required and
schedule, $\Delta$ can vary. In order to satisfy the implicit
restriction imposed by the synchrony hypothesis -- no input event
\textit{from the environment} can be missed -- one needs to calculate
the \textit{Worst Case Reaction Time} (WCRT) and the resultant WCRT
needs to be smaller than the time between any two consecutive events on
any input, else the synchrony hypothesis is violated. Techniques for
calculating the WCRT of synchronous programs~\cite{boldt07} exist and
this analysis is not the focus of this paper. The opposite of the WCRT,
the shortest time for a reaction, is called the \textit{Best Case
  Reaction Time} (BCRT). Formally, let $\{\Delta_1, \Delta_2,\ldots,
\Delta_i\}$, be the set of all possible reaction times for some
CD. Then, $WCRT= max (\Delta_i)$ and $BCRT = min (\Delta_i)$.

We denote the WCRT and BCRT in Figure~\ref{fig:2c} using \texttt{W} and
\texttt{B}, respectively. Figure~\ref{fig:2c} shows a continuously
running physical clock (analogous to a clock in digital hardware), the
numerical annotations on the rising edge of the clock mark the logical
ticks from Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e}
for the numbers show the start and end of the logical ticks,
respectively. The end of a logical tick and the start of the next
logical tick happen together. Figure~\ref{fig:2c} shows the mapping of
the logical time to the physical time. For example, the logical tick
\texttt{1} starts at the first rising edge and completes at the second
rising edge of the physical clock, whereas logical tick \texttt{4}
starts at the $4^{th}$ rising edge, but finishes at the $6^{th}$ rising
edge of the physical clock. This difference can be accounted for by the
Java computations required when signal \texttt{A} is input by the
environment (see Figure~\ref{fig:2a}). This elasticity is inherent (and
elegant) to both Esterel and SystemJ.

### 2014-01-28 correct.tex
Updated correct.tex
Updated things

\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for any
  given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
\label{th:1}
\end{theorem}
\begin{proof}
  We use case analysis and contradiction to prove the above theorem.
  \begin{compactenum}[\hspace{0.25cm} 1.]
  \item Proof for case $T = WCRT$: Assume \mbox{$d \times WCRT < M$},
    then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition, hence,
    $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption to
    hold, $d < \frac{M}{BCRT}$, which contradicts
    Lemma~\ref{lemma:1}. The assumption $d \times WCRT >N$ is trivially
    proven from Lemma~\ref{lemma:1}.
  \item Proof for case $T = BCRT$: Like above, the assumption $d \times
    BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For $d
    \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
    $BCRT \leq WCRT$, by definition, hence, $\frac{N}{BCRT} \geq
    \frac{N}{WCRT}$, which again contradicts Lemma~\ref{lemma:1}.
  \item Proof for case $BCRT < T < WCRT$: For $d \times T < M$ to hold,
    $d < \frac{M}{T}$ should hold. By definition, $T > BCRT$, thus,
    $\frac{M}{BCRT} < \frac{M}{T}$, hence, $d < {M}{BCRT}$ should hold,
    which contradicts Lemma~\ref{lemma:1}. We can prove the other case
    $d \times T \leq N$, using the upper bound from Lemma~\ref{lemma:1}
    similarly.
  \end{compactenum}
\end{proof}

Next, we prove the completeness property.
\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for
  \textbf{all} given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$
  holds.
\end{theorem}
\begin{proof}
  From Theorem~\ref{th:1} case 3.
\end{proof}

@@

\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1}
  \textbf{does not} give a value \texttt{d}, such that for \textbf{all}
  given $BCRT \geq T \geq WCRT$, $M \leq T \times d \leq N$ holds.
\end{theorem}

>>>>

\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for any
  given reaction time $\hj{\Delta}$ \hj{in the range of} $BCRT \leq
  \Delta \leq WCRT$, $M \leq \Delta \times d \leq N$ holds.
\label{th:1}
\end{theorem}
\begin{proof}
  We use case analysis and contradiction to prove the above theorem.
  \begin{compactenum}[\hspace{0.25cm} 1.]
  \item Proof for case $\Delta = WCRT$: Assume \mbox{$d \times WCRT <
      M$}, then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition,
    hence, $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption
    to hold, $d < \frac{M}{BCRT}$, which contradicts
    Lemma~\ref{lemma:1}. The assumption $d \times WCRT >N$ is trivially
    proven from Lemma~\ref{lemma:1}.
  \item Proof for case $\Delta = BCRT$: Like above, the assumption $d
    \times BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For
    $d \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
    $BCRT \leq WCRT$, by definition, hence, $\frac{N}{BCRT} \geq
    \frac{N}{WCRT}$, which again contradicts Lemma~\ref{lemma:1}.
  \item Proof for case $BCRT < \Delta < WCRT$: For $d \times \Delta < M$
    to hold, $d < \frac{M}{\Delta}$ should hold. By definition, $\Delta
    > BCRT$, thus, $\frac{M}{BCRT} < \frac{M}{\Delta}$, hence, $d <
    {M}{BCRT}$ should hold, which contradicts Lemma~\ref{lemma:1}. We
    can prove the other case $d \times \Delta \leq N$, using the upper
    bound from Lemma~\ref{lemma:1} similarly.
  \end{compactenum}
\end{proof}

Next, we prove the completeness property.
\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for
  \textbf{all} given reaction times $\Delta$ in the range of $BCRT \leq
  \Delta \leq WCRT$, $M \leq \Delta \times d \leq N$ holds.
\end{theorem}
\begin{proof}
  From Theorem~\ref{th:1} case 3.
\end{proof}

@@

\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1}
  \textbf{does not} give a value \texttt{d}, such that for \textbf{all}
  given reaction times $T$ in the range of $BCRT > T > WCRT$, $M \leq T
  \times d \leq N$ holds.
\end{theorem}

### 2014-01-28 introduction.tex
Updated things

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about \hj{code
generation} and verification of correctness of functional properties of
the developed programs, but leave nonfunctional properties such as
timing behavior as an implementation detail~\cite{boldt07}. Arguably,
rightly so, because physical time cannot be incorporated without
information about the underlying execution platform.  But, one can still
reason in terms of logical time. These languages support discrete
logical clock rather than a discrete physical (real-time) clock. The
real-time period of the discrete logical clock (also referred to as a
logical tick) is not fixed and it is determined by the responsiveness of
the program to the input signals. Unlike a discrete physical clock,
which has fixed real-time period, the period of the logical clock is
elastic. Although the timing model with logical clock has worked well
for reactive synchronous/GALS languages in designing discrete systems,
there is a need to introduce real-time since often, implementation
models are in the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
capabilities as first class language constructs create additional
opportunity for formally verifying functional and non-functional
requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
\hj{(1) real-time wait does not affect the model of logical time in
	SystemJ program as long as the amount of delay is within boundaries
	that can be determined statically by program analysis,} (2) the
	waiting model relies on the use of relative instead of absolute
	real-time, i.e., a wait in selected time units is counted from the
	currently executing SystemJ statement and (3) these waiting mechanisms
	allow mixing behaviors with real-time features and others with only
	logical time. All these as well as semantics of the wait constructs
	are discussed and illustrated in the rest of the paper. Our
	contributions can be further refined as follows:

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We \hj{clarify} upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for correct by construction code generation
and verification of correctness of functional properties of the
developed programs, but leave nonfunctional properties such as timing
behavior as an implementation detail~\cite{boldt07}. Arguably, rightly
so, because physical time cannot be incorporated without information
about the underlying execution platform.  But, one can still reason in
terms of logical time. These languages support discrete logical clock
rather than a discrete physical (real-time) clock. The real-time period
of the discrete logical clock (also referred to as a logical tick) is
not fixed and it is determined by the responsiveness of the program to
the input signals. Unlike a discrete physical clock, which has fixed
real-time period, the period of the logical clock is elastic. Although
the timing model with logical clock has worked well for reactive
synchronous/GALS languages in designing discrete systems, there is a
need to introduce real-time since often, implementation models are in
the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
capabilities as first class language constructs create additional
opportunity for formally verifying functional and non-functional
requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
\hj{(1) real-time wait does not affect the model of logical time in
  SystemJ program as long as the amount of delay is within boundaries
  that can be determined statically by program analysis, assuming that
  real-time analyzable platforms are available,} (2) the waiting model
relies on the use of relative instead of absolute real-time, i.e., a
wait in selected time units is counted from the currently executing
SystemJ statement and (3) these waiting mechanisms allow mixing
behaviors with real-time features and others with only logical time. All
these as well as semantics of the wait constructs are discussed and
illustrated in the rest of the paper. Our contributions can be further
refined as follows:

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
% \item We consider the property of \textit{non
%     maximal-parallelism}. \textit{Non-maximal parallelism:} There are
%   not always sufficient resources for processes to execute, so
%   effectively processes are allocated and scheduled with different
%   optimization criteria -- reducing computation time, reducing power
%   consumption, etc. This allocation and scheduling is an integral part
%   of any language compilation and needs to be considered when
%   introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We shed light upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

### 2014-01-28 motivation.tex
Updated things

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,wait_exact,suspend}]
{ // Clock-domain 1
while(true) {
  trap(T){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(T); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  wait_exact(1 s); // Wait for 1 second
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; 
      if(#S == 1) counter++; else counter--;}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

@@

The SystemJ program is divided into two \hj{mutually asynchronous}
clock-domains, the first clock-domain continuously displays a green
light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
waits for the user to respond. If the user is able to touch the screen
within the specified time of 50.3 ms to 200.3 ms a positive response is
sent to the second clock-domain. All the SystemJ programming constructs
used to implement this system are described in Table~\ref{tab:1} \hj{and
more details can be found in~\cite{amal10}}. The
\texttt{wait\_inbetween} specification, which models the non exact
delay between 50.3 ms and 200.3 ms is not supported in the current
language and cannot be compiled with the current SystemJ compiler. We
introduce such \texttt{wait\_inbetween} statements in SystemJ. Such
non exact real-time mechanisms are useful in developing systems where
the real-time delay is not known or should not be known a priori.

@@

@@

The timing diagram, Figure~\ref{dd:b}, shows that the
\texttt{wait\_atleast} statement waits for a \textit{at least} 10
seconds, but unlike \texttt{wait\_inbetween} there is no upper
bound. The statement following the \texttt{wait\_atleast} statement is
executed after a finite countable, but unbounded delay. Such a statement
is useful for programming low priority tasks.

@@

>>>>

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,wait_exact,suspend}]
{ // Clock-domain 1
while(true) {
  trap(T){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(T); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; pause; pause; // pass some time before restarting
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; 
      if(#S == 1) counter++; else counter--;}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{d:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

The \texttt{wait\_inbetween (M..N)} statement postpones the program
control flow to move to the proceeding statement for a minimum of $M$
time units and a maximum of $N$ time units, as shown in
Figure~\ref{d:b}.

@@

The SystemJ program is divided into two \hj{mutually asynchronous}
clock-domains, the first clock-domain continuously displays a green
light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
waits for the user to respond. If the user is able to touch the screen
within the specified time of 50.3 ms to 200.3 ms a positive response is
sent to the second clock-domain. All the SystemJ programming constructs
used to implement this system are described in Table~\ref{tab:1} \hj{and
  more details can be found in~\cite{amal10}}. The
\texttt{wait\_inbetween} specification, which models the non exact
postponement between 50.3 ms and 200.3 ms is not supported in the
current language and cannot be compiled with the current SystemJ
compiler. We introduce such \texttt{wait\_inbetween} statements in
SystemJ. Such non exact real-time mechanisms are useful in developing
systems where the real-time postponement is not known or should not be
known a priori.

@@

The \texttt{wait\_atleast (M)} statement is used to postpone the program
control flow from proceeding to the next statment for a minimum of $M$
time units. The maximum postponement is unbounded, but countable
(Figure~\ref{dd:b}).

@@

The timing diagram, Figure~\ref{dd:b}, shows that the
\texttt{wait\_atleast} statement waits for a \textit{at least} 10
seconds, but unlike \texttt{wait\_inbetween} there is no upper
bound. The statement following the \texttt{wait\_atleast} statement is
executed after a finite countable, but unbounded postponement. Such a
statement is useful for programming low priority tasks.

@@

%  LocalWords:  statment
